---
title: 'MIE237'
author: "Neil Montgomery"
date: "2016-01-15"
output: 
  ioslides_presentation: 
    css: '../../../styles.css' 
    widescreen: true 
    transition: 0.001
---
\newcommand{\Var}[1]{\text{Var}\left( #1 \right)}
\newcommand{\E}[1]{E\left( #1 \right)}
\newcommand{\Sample}[1]{#1_1,\ldots,#1_n}
\newcommand{\od}[2]{\overline #1_{#2\cdot}}

# Inference with two independent numerical samples (9.8 and 10.5) continued...

## "Key fact" equal variance version

 Key fact: $$\frac{\left(\od{Y}{1} - \od{Y}{2}\right)
- \left(\mu_1 - \mu_2\right)}{
 S_p\sqrt{\frac{1}{n_1} + \frac{1}{n_2}
}} \sim t_{n_1 + n_2 - 2}$$

## 9.40 continues

```{r}
library(rio)
nitro <- import("Ex09.40.txt")
t.test(nitro$NoNitrogen, nitro$Nitrogen, var.equal = TRUE)
```

## 9.40 with data in "real" form

```{r}
library(tidyr)
nitro_tidy <- gather(nitro, treatment, weight)
```


## Side-by-side boxplots (only n=10 each!)
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
nitro_tidy %>% 
  ggplot(aes(y=weight, x=treatment)) + geom_boxplot()
```

## Summaries

```{r, message=FALSE}
nitro_tidy %>% 
  group_by(treatment) %>% 
  summarize(n(), mean(weight), sd(weight), var(weight))
```

## `t.test` again

```{r}
nitro_tidy %>% 
  t.test(weight ~ treatment, data=., var.equal = TRUE) -> nitro_t
nitro_t
```

Note: `weight ~ treatment` is called a *formula* in R.

## Details for hand calculation

```{r, echo=FALSE}
nitro_tidy %>% 
  group_by(treatment) %>% 
  summarize(n=n(), mean=mean(weight), sd=sd(weight)) -> nitro_summ
nitro_summ
```

$s_p = `r sqrt(sum(nitro_summ$sd^2*(nitro_summ$n - 1))/sum(nitro_summ$n - 2))`$

$\sqrt{\frac{1}{n_1} + \frac{1}{n_2}} = `r sqrt(sum(1/nitro_summ$n))`$

$t_{18, 0.025} = `r qt(0.025, sum(nitro_summ$n-1))`$

$P(t_{18} < `r nitro_t$statistic`) = `r pt(nitro_t$statistic, nitro_t$parameter)`$

## "Key Fact" no equal variance assumption

> Key fact(actually an approximation): $$\frac{\left(\od{Y}{1} - \od{Y}{2}\right)
- \left(\mu_1 - \mu_2\right)}{
\sqrt{\frac{S_1^2}{n_1} + \frac{S_2^2}{n_2}
}} \sim t_\nu$$

> with: $$\nu = \frac{\left(\frac{S^2_1}{n_1} + \frac{S^2_2}{n_2}\right)^2}
{\frac{\left(\frac{S^2_1}{n_1}\right)^2}{n_1-1} +
\frac{\left(\frac{S^2_2}{n_2}\right)^2}{n_2-1}}$$

## 9.40 revisited

```{r}
nitro_tidy %>% 
  t.test(weight ~ treatment, data = .)
```

